Parameter Space Representation of Pareto Front to Explore Hardware-Software Dependencies

### 2014-07-16 algorithm.tex
missing references

\begin{rem}
The innovation score of a region is a measure of how much different
are the points that it adds to Pareto Front compared to the previous
Pareto front. ``High innovation regions'' are the ones in which
more configurations are worth evaluating, because they add many points
on not yet well explored Pareto front areas. Therefore, these regions
are split in the next era (see remark \ref{pers02.rem:smaller_regions})
and more ``crowded'' by simulations. ``No innovation regions''
are the ones that have proved to add few points on Pareto fronts,
probably very near to previous Pareto front points and a Pareto front
point near to previous points is not so much interesting. Therefore
it's not worth evaluating many of their configurations and it is useful
to merge these regions , in order to less densely evaluate them.

ULTERIORI GIUSTIFICAZIONI SULL'IMPORTANZA DI CONSIDERARE LA DIVERSITA'
DELLE SOLUZIONI SI TROVA IN ext12.pag5 E ALTRI.
\end{rem}

\begin{rem}
\label{rem:termination}The algorithm terminates when in the last
$\beta$ iterations not a great ``amount of innovation'' has been
discovered. Therefore it is very probably that carrying on iterating
will not considerably change the Pareto front already formed. It's
not recommended to chose $\beta=1$, because it is possible that the
Pareto fronts $\mathbf{o}\left(\mathscr{P}_{k-1}\right)$ and $\mathbf{o}\left(\mathscr{P}_{k}\right)$
are not very different but next Pareto fronts could be. In other words,
it's unsafe to terminate as soon as the Pareto front doesn't change
enough, passing from an era to the next. It's better taking into account
more than one era.
\end{rem}
!!! IN INGLESE E' PIU' CORRETTO DIRE: ``IT IS POSSIBLE THAT THE PARETO
FRONTS WERE NOT VERY DIFFERENT''? (CON ``WERE'' AL POSTO DI ``ARE'')

>>>>

\begin{rem}
The innovation score of a region is a measure of how much different
are the points that it adds to Pareto Front compared to the previous
Pareto front. ``High innovation regions'' are the ones in which
more configurations are worth evaluating, because they add many points
on not yet well explored Pareto front areas. Therefore, these regions
are split in the next era (see remark \ref{pers02.rem:smaller_regions})
and more ``crowded'' by simulations. ``No innovation regions''
are the ones that have proved to add few points on Pareto fronts,
probably very near to previous Pareto front points and a Pareto front
point near to previous points is not so much interesting. Therefore
it's not worth evaluating many of their configurations and it is useful
to merge these regions, in order to less densely evaluate them.

%ULTERIORI GIUSTIFICAZIONI SULL'IMPORTANZA DI CONSIDERARE LA DIVERSITA'
%DELLE SOLUZIONI SI TROVA IN ext12.pag5 E ALTRI.

\begin{rem}
\label{rem:termination}The algorithm terminates when in the last
$\beta$ iterations not a great ``amount of innovation'' has been
discovered. Therefore it is very probably that carrying on iterating
will not considerably change the Pareto front already formed. It's
not recommended to chose $\beta=1$, because it is possible that the
Pareto fronts $\mathbf{o}\left(\mathscr{P}_{k-1}\right)$ and $\mathbf{o}\left(\mathscr{P}_{k}\right)$
are not very different but next Pareto fronts could be. In other words,
it's unsafe to terminate as soon as the Pareto front doesn't change
enough between two consecutive eras, thus a strategy taking into
account the history of the Pareto fronts has been adopted in the
proposed approach.
\end{rem}
%!!! IN INGLESE E' PIU' CORRETTO DIRE: ``IT IS POSSIBLE THAT THE PARETO
%FRONTS WERE NOT VERY DIFFERENT''? (CON ``WERE'' AL POSTO DI ``ARE'')

### 2014-07-16 experiments.tex
missing references

\section{A Case Study: VLIW Hardware/Software Design Space}
\label{sec:ee}
In order to evaluate the behavior of the proposed approach
we will be using a parameterized Very Long Instruction
Word (VLIW) architecture~\cite{kathail_tr00} as testbed. The same philosophy
behind a VLIW system makes this a natural choice for several reasons:
\begin{itemize}
\item{Multi-Objective trade-offs}: VLIW architectures let designers
easily trade between power, energy and performances~\cite{}, leading
to Pareto sets whose extension is an ideal ground for our testing
purposes focused on parameter-space space representation of Paretos.
\item{Compiler awareness}: Instruction Level Parallelism (ILP)
constitutes the key feature to enable the multi-objective trade offs
of the point above. It is statically obtained by the compiler which
performs code trasformations being aware of the underlying hardware
configuration.  This tight hardware/software dependence, makes the
VLIW scenario well suitable for investigating hardly predictable
hardware/software interactions, especially when compared to
architectures that, in order to preserve binary compatibility,
separate software generation processes from the hardware that
will execute it.

System parameters can be classified in three main categories:
\emph{processor}, \emph{memory sub-system} and \emph{compiler}. 
The first is directly related to the VLIW core and includes the size
of the register files and the number of functional
units for each type. In particular, five different types of register
files are supported:
GPR (32-bit registers for integers), FPR (64-bit registers for
floating point values) PR (1-bit registers used to store the Boolean
values of predicated instructions), BTR (64-bit registers containing
information about possible future branches) and CR (32-bit control
registers containing information about the internal state of the
processor). The functional units involved are: \emph{Integer units},
\emph{floating point units}, \emph{memory units} (associated with
load/store operations) and \emph{branch units} (associated with branch
operations). With respect to the memory sub-system, the parameters
include \emph{size}, \emph{associativity} and
\emph{block size} for each of the three caches: First-level data cache
(L1D), first-level instruction cache (L1I) and second-level unified
cache (L2U).

@@

@@

The results presented in the next section compare PS against a widely
adpted Multi-objective Genetic Algorithm~\cite{}. At best of our
knowledge, genetic approaches
still represent the best compromise between efficiency (i.e. time
required for exploration) and accuracy of the reported Paret Sets.
Note that for the reasons described above, mono-objective approaches have been
explicitly excluded from this analysis. Same decision has been taken
for other approaches (e.g. dependecy based~\cite{dep}) that make use of some a-priori knowledge about the
role and meaning of each parameter, since our
aim is evaluate the capacity of focusing on interesting
parameter regions without any external judgement or help based on
euristics.  Finally, it should be pointed
out that in this introductory work we will explicitly use the TODO
variant of MOGA, remainding to works like~\cite{} for a detailed
analysis on how it differs and what could be expected from similar multi-objective strategies.

The class of benchmark being considered is representative of some common
frequently running application kernels in an embedded multimedia
environment. Table~\ref{tab:bench} shows the set of applications
chosen along with a brief description.
\begin{table}
	\centering
	\caption{Benchmarks}
	\label{tab:bench}
	\begin{tabular}{ll}
	\hline
	\multicolumn{1}{c}{Benchmark} & \multicolumn{1}{c}{Application} \\
	\hline
	Alloca\_test & Memory array allocation test \\
	Bmm & Matrices multiplication and elements sum \\
	Fir\_int & Finite impulse response \\
	Mm & Floating point matrices multiplication \\
	Sha256 & Cryptocurrency header Hashing \\
	Wave & Wavefront computation \\
	\hline
	\end{tabular}
\end{table}
%------------------------------------------------------------------------------

@@

In order to evaluate the Paretos from a quantitative point-of-view we
will consider two metrics, namely, the \emph{variation range} and the
\emph{average normalised absolute dispersion error}. For a given
objective, the variation range represents the ratio between the
maximum and the minimum value observed for that objective. A
comparison between the variation range for different benchmarks
between a FCP and a VCP exploration for both power dissipation and
execution time is shown in Fig.~\ref{fig:variation_range}. As it can
be observed, the VCP exploration provides solutions which fall on a
range that is, on average, 23\% and 40\% wider than that provided by a
FCP exploration for power dissipation and execution time,
respectively.
\begin{table}
  \centering
  \begin{tabular}{c}
    \includegraphics[width=0.4\textwidth]{pictures/variation_power.eps} \\
    (a) \\
    \includegraphics[width=0.4\textwidth]{pictures/variation_etime.eps}
    (b) \\
    \includegraphics[width=0.4\textwidth]{pictures/dispersion.eps} \\
    (c) 
  \end{tabular}
  \caption{(a)(b) Variation range for different benchmarks between PS and GA
  and average dispersion absolute dispersion (c)}
  \label{fig:dispersion}
\end{table}

>>>>

\section{A Case Study: VLIW Hardware/Software Design Space}
\label{sec:ee}
\subsection{Motivation}
In order to evaluate the behavior of the proposed approach
we will be using a parameterized Very Long Instruction
Word (VLIW) architecture~\cite{kathail_tr00} as testbed. The same philosophy
behind a VLIW system makes this a natural choice for several reasons:
\begin{itemize}
\item{Multi-Objective trade-offs}: VLIW architectures let designers
easily trade between power, energy and performances~\cite{}, resulting
in Pareto sets whose extension is an ideal ground for our testing
purposes focused on parameter-space space representation of Paretos.
\item{Compiler awareness}: Instruction Level Parallelism (ILP)
constitutes the key feature to enable the multi-objective trade offs
of the point above. It is statically obtained by the compiler which
performs code trasformations being aware of the underlying hardware
configuration.  This tight hardware/software dependence, makes the
VLIW scenario well suitable for investigating hardly predictable
hardware/software interactions, especially when compared to
architectures that, in order to preserve binary compatibility,
separate software generation processes from the hardware that
will execute it.
\end{itemize}

\subsection{Parameter Space}
System parameters can be classified in three main categories:
\emph{processor}, \emph{memory sub-system} and \emph{compiler}. 
The first is directly related to the VLIW core and includes the size
of the register files and the number of functional
units for each type. In particular, five different types of register
files are supported:
GPR (32-bit registers for integers), FPR (64-bit registers for
floating point values) PR (1-bit registers used to store the Boolean
values of predicated instructions), BTR (64-bit registers containing
information about possible future branches) and CR (32-bit control
registers containing information about the internal state of the
processor). The functional units involved are: \emph{Integer units},
\emph{floating point units}, \emph{memory units} (associated with
load/store operations) and \emph{branch units} (associated with branch
operations). With respect to the memory sub-system, the parameters
include \emph{size}, \emph{associativity} and
\emph{block size} for each of the three caches: First-level data cache
(L1D), first-level instruction cache (L1I) and second-level unified
cache (L2U).

@@

As regards the class of benchmark being considered, a set applications representative of some common
frequently running code kernels in an embedded multimedia
environment has been chosen. Table~\ref{tab:bench} shows the set of applications
along with a brief description.
\begin{table}
	\centering
	\caption{Benchmarks}
	\label{tab:bench}
	\begin{tabular}{ll}
	\hline
	\multicolumn{1}{c}{Benchmark} & \multicolumn{1}{c}{Application} \\
	\hline
	Alloca\_test & Memory array allocation test \\
	Bmm & Matrices multiplication and elements sum \\
	Fir\_int & Finite impulse response \\
	Mm & Floating point matrices multiplication \\
	Sha256 & Cryptocurrency header Hashing \\
	Wave & Wavefront computation \\
	\hline
	\end{tabular}
\end{table}

@@

The results presented in the next section compare PS against one
widely adopted Multi-objective Genetic
Algorithm~\cite{knowles_techrep06}. At best of our knowledge, genetic
approaches in general still represent the best compromise between
efficiency (i.e. time required for exploration) and accuracy of the
reported Paret Sets.  Note that for the reasons described above,
mono-objective approaches have been explicitly excluded from this
analysis. Same decision has been taken for other approaches (e.g.
dependecy based~\cite{givargis_tvlsi02}) that make use of some
a-priori knowledge about the role and meaning of each parameter, since
our aim is evaluate the capacity of focusing on interesting parameter
regions without any external judgement or help based on euristics.
Finally, it should be pointed out that in this introductory work we
will explicitly use the SPEA~\cite{zitzler_eurogen01} variant of MOGA,
remainding to works like~\cite{zitzler_ec00}~\cite{zitzler_tec03} for
a detailed analysis on how it differs and what could be expected from
similar multi-objective strategies.

%------------------------------------------------------------------------------

@@

In order to evaluate the Paretos from a quantitative point-of-view we
will consider two metrics, namely, the \emph{variation range} and the
\emph{average normalised absolute dispersion error}. For a given
objective, the variation range represents the ratio between the
maximum and the minimum value observed for that objective. A
comparison between the variation range for different benchmarks
between a FCP and a VCP exploration for both power dissipation and
execution time is shown in Fig.~\ref{fig:dispersion}(a) and (b)
respectively. As it can
be observed, the VCP exploration provides solutions which fall on a
range that is, on average, 23\% and 40\% wider than that provided by a
FCP exploration for power dissipation and execution time,
respectively.
\begin{table}
  \centering
  \begin{tabular}{c}
    \includegraphics[width=0.4\textwidth]{pictures/variation_power.eps} \\
    (a) \\
    \includegraphics[width=0.4\textwidth]{pictures/variation_etime.eps}
    (b) \\
    \includegraphics[width=0.4\textwidth]{pictures/dispersion.eps} \\
    (c) 
  \end{tabular}
  \caption{(a)(b) Variation range for different benchmarks between PS and GA
  and average dispersion absolute dispersion (c)}
  \label{fig:dispersion}
\end{table}
